Certified Knowledge Compilation \\ with Application to Formally Verified Model Counting

### 2024-01-05 doc/JAIR/jair-paper.tex
Minor polishing

The forward implication proof of Figure~\ref{fig:eg4:all}(E) was
generated by the structural approach.  Working from step 36 backward,
we can see that steps 35 and 36 complete the call to
$\validate(\nodes_{10}, \emptyset, \phi_I)$.  This call used $x_1$ as
the splitting variable, first calling $\validate(\nodep_{8},
\{\obar{x}_1\}, \emptyset, \phi_I)$, which completed with step 29, and
$\validate(\nodep_{9}, \{x_1\}, \phi_I)$, which completed with step
34.  We see that each of these calls required separate traversals of
nodes $\nodes_7$, $\nodep_6$, and $\nodep_5$, with the former yielding
proof steps 25--27 and the latter yielding proof steps 30--32.  This
demonstrates how our simple formulation of $\validate$ effectively
expands the graph into a tree.  This shortcoming is avoided by the use of lemmas, as is described in Section~\ref{sect:lemma}.

@@

Suppose for every clause $C_j \in \gamma_{\nodeu}$ that does not correspond to
an input clause, we generate a synthetic clause $C'_j$ with guard literal
$\obar{v}_j$, for $1 \leq j \leq m$.  Let $\gamma'_{\nodeu}$ be the formula where each clause $C_j$ is replaced by synthetic clause $C'_j$,
while input clauses in $\gamma_{\nodeu}$ are left unchanged.
Let $\lassign = \{ \obar{v}_1, \obar{v}_2, \ldots, \obar{v}_m \}$.
Invoking $\validate(\nodeu, \lassign, \gamma'_{\nodeu})$
 will then prove a lemma, given by the target clause
 $u \lor v_1 \lor v_2 \lor \cdots \lor v_m$,
 showing that any total assignment $\assign$ that activates the synthetic clauses will cause $u$ to be assigned $1$.
 More precisely, given assignment $\assign$ and its extension $\eassign$, if $\eassign(v_j) = 0$ for every guard literal $\obar{v}_j$, then $\eassign(u) = 1$.

@@

Observe that the lemma structure can be hierarchical, since a shared
subgraph may contain nodes that are themselves roots of shared
subgraphs.  Nonetheless, the principles described allow the
definition, proof, and applications of a lemma for each shared node in
the graph.  For any node $\nodeu$, the first call to
$\validate(\nodeu, \passign, \psi)$ may require further recursion,
but any subsequent call can simply reuse the lemma proved by the first call.

>>>>

As examples, the forward implication proof of Figure~\ref{fig:eg4:all}(E) was
generated by the structural approach.  Working from step 36 backward,
we can see that steps 35 and 36 complete the call to
$\validate(\nodes_{10}, \emptyset, \phi_I)$.  This call used $x_1$ as
the splitting variable, first calling $\validate(\nodep_{8},
\{\obar{x}_1\}, \emptyset, \phi_I)$, which completed with step 29, and
$\validate(\nodep_{9}, \{x_1\}, \phi_I)$, which completed with step
34.  We see that each of these calls required separate traversals of
nodes $\nodes_7$, $\nodep_6$, and $\nodep_5$, with the former yielding
proof steps 25--27 and the latter yielding proof steps 30--32.  This
demonstrates how our simple formulation of $\validate$ effectively
expands the graph into a tree.  This shortcoming is avoided by the use of lemmas, as is described in Section~\ref{sect:lemma}.

@@

Suppose for every clause $C_j \in \gamma_{\nodeu}$ that does not correspond to
an input clause, we generate a synthetic clause $C'_j$ with guard literal
$\obar{v}_j$, for $1 \leq j \leq m$.  Let $\gamma'_{\nodeu}$ be the formula where each clause $C_j$ is replaced by synthetic clause $C'_j$,
while input clauses in $\gamma_{\nodeu}$ are left unchanged.
Let $\lassign = \{ \obar{v}_1, \obar{v}_2, \ldots, \obar{v}_m \}$.
Invoking $\validate(\nodeu, \lassign, \gamma'_{\nodeu})$
 will then prove a lemma, given by the target clause
 $u \lor v_1 \lor v_2 \lor \cdots \lor v_m$,
 showing that any total assignment $\assign$ that activates the synthetic clauses will, when cause $u$ to be assigned $1$.
 More precisely, given assignment $\assign$ and its extension $\eassign$, if $\eassign(v_j) = 0$ for every guard literal $\obar{v}_j$, then $\eassign(u) = 1$.

@@

Observe that the lemma structure can be hierarchical, since a shared
subgraph may contain nodes that are themselves roots of shared
subgraphs.  Even then, the principles described allow the
definition, proof, and applications of a lemma for each shared node in
the graph.  For any node $\nodeu$, the first call to
$\validate(\nodeu, \passign, \psi)$ may require further recursion,
but any subsequent call can simply reuse the lemma proved by the first call.
